---
title: "Data Preprocessing"
author: "Mason Ogden"
date: "6/1/2021"
output: html_document
---

[Link](https://www.kaggle.com/c/shelter-animal-outcomes/data) to Kaggle competition

**Packages**

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(magrittr)
library(lubridate)
library(tidymodels)

set.seed(143)
```

**Data**

```{r, message = FALSE}
dir <- here::here()
dataset_dir <- dir %>%
  paste0("/datasets")

df <- dataset_dir %>%
  paste0("/train.csv") %>%
  read_csv() %>%
  rename(animal_id = AnimalID, animal_name = Name, outcome_datetime = DateTime, 
         outcome = OutcomeType, outcome_subtype = OutcomeSubtype, 
         animal_type = AnimalType, sex = SexuponOutcome, 
         age = AgeuponOutcome, breed = Breed, color = Color) %>%
  relocate(outcome, .after = "animal_id")

test_df <- dataset_dir %>%
  paste0("/test.csv") %>%
  read_csv() %>%
  rename(animal_id = ID, animal_name = Name, outcome_datetime = DateTime, 
         animal_type = AnimalType, sex = SexuponOutcome, 
         age = AgeuponOutcome, breed = Breed, color = Color) %>%
  mutate(animal_id = as.character(animal_id))

df_small <- df %>% 
  slice_sample(n = 1000)
```

Observations:
  1. ~~ID column that needs to be dropped for prediction~~
  2. ~~Lots of missing values for 'name', but probably won't be used for prediction~~
    + unless you want to do some kind of sentiment analysis with transfer learning
  3. ~~Not sure how to use/if I should use the 'OutcomeSubtype' variable, seems like a more specific version of the response variable 'Outcome'. ~~
    + removed
  4. ~~Probably a ton of unique values in the 'breed' column~~
    + I converted it to a simple binary 'mix/not mix'
  5. ~~There's a datetime column that probably contains some useful information, like when the outcome happened (month, time of day, year, etc...), holiday.~~ 
  6. ~~'Age' column needs to be converted to numeric.~~
  
### Preprocessing

**Converting age to numeric**

```{r}
age_to_years_numeric <- function(age_char) {
  case_when(
    # if measured in years
    str_detect(age_char, pattern = "year") ~ as.double(str_extract(age_char, pattern= "\\d+")),
    # if measured in months
    str_detect(age_char, pattern = "month") ~ as.double(str_extract(age_char, pattern = "\\d+")) / 12,
    # if measured in weeks
    str_detect(age_char, pattern = "week") ~ as.double(str_extract(age_char, pattern = "\\d+")) / 52.1429,
    # if measured in days
    str_detect(age_char, pattern = "day") ~ as.double(str_extract(age_char, pattern = "\\d+")) / 365
  )
}
```

**Simplifying the 'breed' variable**
  
```{r}
collapse_breeds <- function(breed_vec) {
  ifelse(str_detect(breed_vec, pattern = "Mix|\\/"), "Mix", breed_vec)
}
```

**Reducing number of unique animal colors**

```{r}
collapse_colors <- function(colors_fac) {
  colors_vec <- as.character(colors_fac)
  
  case_when(
    str_detect(colors_vec, pattern = "Tabby") ~ "Tabby",
    str_detect(colors_vec, pattern = "Point") ~ "Point",
    str_detect(colors_vec, pattern = "Brindle") ~ "Brindle",
    str_detect(colors_vec, pattern = "Calico") ~ "Calico",
    str_detect(colors_vec, pattern = "Merle") ~ "Merle",
    str_detect(colors_vec, pattern = "Tricolor") ~ "Tricolor",
    str_detect(colors_vec, pattern = "Sable") ~ "Sable",
    str_detect(colors_vec, pattern = "Blue Tick") |  str_detect(colors_vec, pattern = "Blue Cream") ~ "Blue",
    str_detect(colors_vec, pattern = "Black Smoke") ~ "Black",
    str_detect(colors_vec, pattern = "/") ~ "Multiple colors",
    str_detect(colors_vec, pattern = "Tortie") | str_detect(colors_vec, pattern = "Torbie") ~ "Tortie/Torbie",
    TRUE ~ colors_vec
  )
}
```

**Preprocessing Recipe**

Does everything below but in one big step:

```{r}
preproc_rec <- recipe(outcome ~ ., data = df_small) %>%
  # set 'animal_id' as the identifier variable, not to be used in prediction
  update_role(animal_id, new_role = "id") %>%
  # data preprocessing
  step_mutate(
    # convert age to numeric
    age = age_to_years_numeric(age),
    # add month variable (as a factor)
    outcome_month = factor(month(outcome_datetime)),
    # add year variable (as a factor)
    outcome_year = factor(year(outcome_datetime)),
    # reduce number of unique breeds
    mix = factor(collapse_breeds(breed)),
    # reduce number of unique animal colors
    color = factor(collapse_colors(color))
    ) %>%
  # add indicator variable for if date is a holiday
  step_holiday(outcome_datetime,
               holidays = c("ChristmasEve", "NewYearsDay", "USChristmasDay",
                            "USThanksgivingDay")) %>%
  # remove columns that are no longer useful
  step_rm(outcome_datetime, breed, animal_name, outcome_subtype)
```

**Applying Recipe**

```{r}
# fit to training data
preproc_prepped <- preproc_rec %>% prep(df)

# apply to training data
preproc_train <- preproc_prepped %>% bake(df)

# apply to test data
preproc_test <- preproc_prepped %>% bake(test_df) %>%
  mutate(animal_id = as.character(1:nrow(test_df)))

preproc_test %>% head()
```


### Outputting Preprocessed Data

```{r}
preproc_dir <- dir %>%
  paste0("/preprocessed_data")

preproc_train %>%
  write_csv(paste0(preproc_dir, "/train_preprocessed.csv"))

preproc_test %>%
  write_csv(paste0(preproc_dir, "/test_preprocessed.csv"))
```

