---
title: "Modeling"
author: "Mason Ogden"
date: "6/1/2021"
output: html_document
---

[Link](https://www.kaggle.com/c/shelter-animal-outcomes/data) to Kaggle competition

Questions: 

  + OK to just use every classification model included in Tidymodels?

**Packages**

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(magrittr)
library(lubridate)
library(tidymodels)
library(baguette)
library(plsmod)
library(rules)
library(discrim)

set.seed(143)
```

**Data**

```{r, message = FALSE, warning = FALSE}
dir <- here::here()

preproc_dir <- dir %>%
  paste0("/preprocessed_data")

train <- preproc_dir %>%
  paste0("/train_preprocessed.csv") %>%
  read_csv()

test <- preproc_dir %>%
  paste0("/test_preprocessed.csv") %>%
  read_csv()
```

**Approaches**

[Link](https://www.tidymodels.org/find/parsnip/) to list of Tidymodels models

1. Bagged MARS model
  + engine: "earth"
  + bag_mars()
  
2. Bagged Decision Tree
  + engine: "rpart"
  + bag_tree()
  
3. Boosted Tree
  + engine: "xgboost"
  + boost_tree()
  
4. Decision Tree
  + engine: "rpart"
  + decision_tree()
  
5. MARS
  + "earth"
  + mars()
  
6. Multi-layer Perceptron
  + just a single layer neural network
  + engine: "keras"
  + mlp()
  
7. KNN
  + engine: "kknn"
  + nearest_neighbor()
  
8. Partial Least Squares (PLS)
  + engine: "mixOmics"
  + pls()
  
9. Random Forest
  + engine: "randomForest"
  + rand_forest()
  
10. RuleFit Models
  + engine: "xrf"
  + rule_fit()
  
11. Polynomial SVM
  + engine: "kernlab"
  + svm_poly()

12. RBF SVM
  + engine: "kernlab"
  + svm_rbf()
  
13. C5.0 Rule-based Classification
  + engine: "C5.0"
  + C5_rules()
  
14. Flexible Discriminant Model
  + engine: "earth"
  + discrim_flexible()
  
15. LDA
  + engine: "MASS"
  + discrim_linear()
  
16. Regularized LDA
  + engine: "klaR"
  + discrim_regularized()
  
17. Multinomial Regression
  + engine: "glmnet"
  + multinom_reg()
  
18. Naive Bayes
  + engine: "klaR"
  + naive_bayes()
  
19. Neural Network
  + made using keras from scratch
  + look up architectures for classification